\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2018-11-26 write_up/mixfishsim_ecolmod.tex
Text revisions - clarifying and updating text

\replaced[id = PD]{If fisheries are to avoid unwanted catch}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions, movement and how fishers interact with different fish
populations. This reflects the fact that data on fish location at high temporal
and spatial resolutions is expensive and difficult to collect. Proxies inferred
from either scientific surveys or commercial catch data are often used to model
distributions, often with data at limited spatial and temporal resolution. \\

@@

We simulate 50 years of fishing and use the results from the fisheries catch to
draw inference on the underlying population structures. We compare this
inference to i) a simulated fixed-site sampling design commonly used for
fisheries monitoring purposes, and ii) the true underlying population
structures input to the simulation, to establish the potential and limitations
of fishery-dependent data - an inherently biased sampling method due to
fisher's targeting- to provide a robust picture of spatiotemporal
distributions. Finally, we simulate an area closure based on areas defined from
commercial the known ("real-population") distribution, commercial catch data
and survey data at different temporal and spatial resolutions and assess their
effectiveness on reducing catches of a fish population. \\

@@

In order to understand \replaced[id=PD]{the consequences of using}{challenges
	that face} VMS-linked landings to draw inference on the underlying
population structure we develop a simulation model where population dynamics
are highly-resolved in space and time\replaced[id=PD]{. Being}{ and are}
known \added[id=PD]{directly} rather than inferred from sampling or commercial
catches \added[id=PD]{, we can use the population model to evaluate how inference from
	fisheries-dependent and fisheries independent sampling relates to the
	real population structure}. \replaced[id=PD]{In our model system
	p}{P}opulation movement is driven by random (diffusive) and directed
(advective) processes and we incorporate characterisation of a number of
different fisheries \added[id=PD]{dynamics} exploiting four fish populations
with different spatial and population demographics.\\

@@

\begin{enumerate}
	\item to understand how sampling-derived data reflects the underlying
		population structures. We compare at different spatial and
		temporal aggregations of data the real population to: 
		\begin{enumerate}
			\item the inferred population from a stratified fixed-site sampling survey design
				commonly used for fisheries monitoring purposes, otherwise know as a
				fisheries-independent survey,
			\item the inferred population from our fishery-dependent model which includes
				fishery-induced sampling dynamics.\\
		\end{enumerate}
	\item to understand the impact of data aggregation and source on
		spatial fisheries management measures we simulate a fishery closure to
		protect a species based on different spatial and temporal data
		aggregations:
		\begin{enumerate}
			\item as if the real spatial population structure were
				known,
			\item the fishery-independent inferred population
				structure
			\item the fishery-dependent inferred population
				structure
		\end{enumerate}
		We evaluate the theoretical "benefit" to the population of the
		closure(s), and effect on the other three
		populations\todo{\added[id=JJ]{If the paper has two goals this
				should be clear from the start, but may be
				better over 2 MSs}\added[id=PD]{I would like to
				keep both parts, but have made clearer in how
				its set out. The closure scenarios form
				validation of the data aggregation, rather than
				effectiveness of the closures themselves - so
				its a continuation of the same question in my
				eyes}}\deleted[id=PD]{.  Further, we extend our
			analysis to a range of spatial and temporal scales to
			assess the impact of these processes on the success of
			the management measure.}\\
\end{enumerate}

@@

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics (fishing and natural mortality,
growth) operate on a daily time-step\replaced[id=PD]{, while p}{. P}opulation
movement occurs on a weekly time-step\replaced[id=PD]{. R}{, while r}ecruitment
\replaced[id=PD]{takes place}{occurs} periodcally each year for a set time
\replaced[id=PD]{duration}{period (e.g. 3 weeks)} at at specified point
individual to a species\deleted[id=PD]{.}\added[id=PD]{, while the fishing
	module operates on a tow-by-tow basis (i.e. multiple events a day)}.
The simulation framework is implemented in the statistical software package R
\citep{RCoreTeam2017} \replaced[id=PD]{and}{;} available as an R package from
the authors github (\url{www.github.com/pdolder/MixFishSim}).\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Overview schematic  
\begin{figure}[!ht]
	\hspace{-4em}
\begin{tikzpicture}[%
    >=triangle 60,              % Nice arrows; your taste may be different
    start chain=going below,    % General flow is top-to-bottom
    node distance=8mm and 60mm, % Global setup of box spacing
    every join/.style={norm},   % Default linetype for connecting boxes
    scale=0.9,
    every node/.style={scale=0.8}]                 
    \label{fig:model}
    % ------------------------------------------------- 
% A few box styles 
% <on chain> *and* <on grid> reduce the need for manual relative
% positioning of nodes
\tikzset{
  base/.style={draw, on chain, on grid, align=center, minimum height=4ex},
  proc/.style={base, rectangle, text width=8em},
  test/.style={base, rectangle, text width=5em},
  term/.style={proc, rounded corners},
  % coord node style is used for placing corners of connecting lines
  coord/.style={coordinate, on chain, on grid, node distance=6mm and 25mm},
  % nmark node style is used for coordinate debugging marks
  nmark/.style={draw, cyan, circle, font={\sffamily\bfseries}},
  % -------------------------------------------------
  % Connector line styles for different parts of the diagram
  norm/.style={->, draw, lcnorm},
  free/.style={->, draw, lcfree},
  cong/.style={->, draw, lccong},
  it/.style={font={\small\itshape}}
}
% -------------------------------------------------
% Start by placing the nodes in the middle
\node [proc, densely dotted, it] (p0) {start};
% Use join to connect a node to the previous one 
\node [term, join, fill=lcnorm!25]      {Choose random location to fish};
\node [term, join, fill=lcfree!25] (p1){Sample catch};
\node [test, densely dotted] (t1) {if $Rec_{p}[t]$};
\node [test, join, densely dotted] (t2) {if $wk[t] \neq  wk[t-1]$};
\node [test, join, densely dotted] (wk) {if $day[t] \neq day[t-1]$};
\node [test, join, densely dotted] (t3) {if $trip[t] \neq trip[t-1] \ | \ PK[t]$};
\node [term, join, fill=lcnorm!25]  (p2) {Correlated Random Walk};
\node [term, join, fill=lcnorm!25]  (p3) {Choose new fishing location};
\node [proc, densely dotted, join, it] (p7) {end};

@@

% -------------------------------------------------
% A last flourish which breaks all the rules
\draw [->,purple, dotted, thick, shorten >=1mm]
  (p3.south) -- ++(5mm,-3mm)  -- ++(87mm,0mm) 
  |- node [black, near end, yshift=1.75em, it]
    {} ($(p1.north) + (0,0.3)$);
% -------------------------------------------------
\end{tikzpicture}
\caption{Overview Schematic of simulation model. The blue boxes indicate fleet
	dynamics processes, the green boxes population dynamics processes while
	the white boxes are the timesteps at which processes occur; $Rec$
	is a recruitment period for population $p$, $t$ = tow, $tmax$ is the
	total number of tows, $wk$ is a weekly timestep, $day$ is a day
	timestep, $trip$ is a trip time step.}
\label{fig:overview}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen as to discretise continuous population processes
	on a biologically relevent and computationally tractable timescale.}
\replaced[id=PD]{Under the popualation dynamics module}{Here,} population
biomass growth and depletion for pre-recruits and \deleted[id=PD]{fish}
recruited \added[id=PD]{fish} \deleted[id=PD]{to the fishery} are modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment:

\begin{equation*}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation*}

where $\rho$ is Brody's coefficient, shown to be approximately equal to
$exp(-K)$, where $K$ is the growth rate from a von bertalanffy logistic growth
model \citep{Schnute1985}. $Wt_{R-1}$ is the weight of fish prior to
recruitment, while $Wt_{R}$ is the recruited weight. $\alpha_{d}$ represents
the proportion of fish recruited during that day for the year, while
$R_{\tilde{c,y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:

\begin{equation*}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})}) *
B_{c,d}
\end{equation*}

where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. \todo{\added[id=CM]{[link $F$ to
	effort and catchability - as I think we have F as an emergent property
	of the fleets rather than something we solve for (I could be wrong
	though!) - catch for a vessel is a product of catchability and biomass,
	i.e. C = qB, but this catch is summed to solve for F. So its both
	really]}}\\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{mixfishsim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}:

\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * B_{c,d})}{(\beta + B_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum, $B$ current SSB and
$\sigma^2$ the variability in the recruitment due to stochastic
processes. \\

or a stochastic Ricker form \citep{Ricker1954}:

\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}

where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations\replaced[id=JJ]{. An}{,
	with an} advection-diffusion process \deleted[id=JJ]{to}
control\added[id=JJ]{led} \deleted[id=JJ]{how the} population\deleted[id=JJ]{s}
move\replaced[id=JJ]{ment}{d} over time with a moving temperature covariate to
capture temporal dependencies\todo{\added[id=JJ]{What have a temperature
		covariate? Could just use time}\added[id=PD]{Was intended as
		some biological meaning - species thermal tolerances load onto
		the temperature effect}}. \deleted[id=JJ]{This was intended to
	balance realism in population movement, capturing the main directed and
	random processes, and practicality of modelling the population rather
	than individual fish.} \\

@@

The covariance structure affects the smoothness of the surfaces which the
process generates\replaced[id=PD]{;}{, and} we used the \textit{Matérn}
\deleted[id=PD]{family of} covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{as}{one where} the correlation strength weakens the further
the distance apart \deleted[id=PD]{(i.e. the correlation between $S(x)$ and
	$S(x')$ decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{The Matérn covariance structure models the spatial
	autocorrelation observed with animal distributions \citep{Tobler1970,
		F.Dormann2007}} \replaced[id=PD]{and}{The \textit{Matérn}
	correlation} is a two-parameter family where:

\begin{center}
	$\rho(u) = \{2^{\kappa -
		1}\Gamma{\kappa}\}^{-1}(u/\phi)^{\kappa}K_{\kappa}(u/\phi)$
\end{center}

$K_{\kappa}(.)$ is a modified Bessel function of order $\kappa$, $\phi >
0$ is a scale parameter with the dimensions of distance, and $\kappa > 0$,
called the order, is a shape parameter which determines the smoothness of the
underlying process. \\

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, implementing different parameter settings to affect the
patchiness of the populations. Each population \replaced[id=PD]{was}{is}
initialised at a single location, and subsequently move\replaced[id=PD]{ed}{s}
according to a probabilistic distrib\added[id=PD]{u}tion based on habitat
suitability, temperature and distance from current cell\replaced[id=PD]{:}{.}

\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}

Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$.\\

During specified weeks of the year, the habitat quality
\replaced[id=PD]{was}{is} modified for \added[id=PD]{user-defined} spawning
habitat\deleted[id=PD]{s}, \replaced[id=PD]{resulting in}{meaning} each
population \replaced[id=PD]{had}{has} a concentrated area where spawning takes
place and the population move\replaced[id=PD]{d}{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to
spawning.\todo{\added[id=JM]{What does it mean concisely? Areas are assigned?}
\added[id=PD]{Yes, the areas are pre-defined - I have amended to reflect and
	tried to clarify}} \\

The temperature field \replaced[id=PD]{was}{is} simulated to be on a gradient
from a South-Westerly to North-Easterly direction, with temperature in each
cell changing gradually on a week-by-week basis so that initially high
temperature areas cycle\added[id=PD]{d} to lower temperatures and low
temperature areas vice versa. Each population $p$ \replaced[id=PD]{wa}{i}s
assigned a thermal tolerance with mean, $\mu\added[id=PD]{_{p}}$ and variance,
$\sigma^2\added[id=PD]{_{p}}$ so that each cell and population temperature
suitability is defined that:

\begin{equation}
	Tol_{c, p\added[id=PD]{, wk}} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}

Where $Tol_{c, p, \added[id=PD]{wk}}$ is the tolerance of population $p$ in
cell $c$ \added[id=PD]{in week $wk$}, $T_{c, \added[id=PD]{wk}}$ is the
temperature in the cell \added[id=PD]{given the week} and
$\mu\added[id=PD]{_{p}}$ and $\sigma^2\added[id=PD]{_{p}}$ the mean and
standard deviation of the population temperature tolerance. \\

\added[id=PD]{The final process resulted in independent populations structure
	and movement patterns, with population movement occuring on a weekly
	basis. This process approximated the demographic shifts in fish
	populations throughout a year while maintaining seasonal patterns for
	spawning.}

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, \textit{Q}, and a population specific efficiency,
${Q_{p}}$.  Thus, the product of these parameters affect\replaced[id=PD]{s}{ed}
the overall catch rates for the fleet and the preferential targeting of one
population over another. This, in combination with the parameter choice for the
step-function \added[id=PD]{defined below} (as well as some randomness from the
exploratory fishing process) determine\replaced[id=PD]{d}{s} the preference of
fishing locations for the fleet.  All species prices
\replaced[id=PD]{were}{are} kept the same, across fleets\replaced[id=PD]{and
seasons}{, though can be made to vary seasonally}.

@@

\subsection{Population parameterisation}

>>>>

\replaced[id = PD]{If fisheries are to avoid unwanted catch}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions, movement and how fishers interact with different fish
populations. This reflects the fact that data on fish location at high temporal
and spatial resolutions is expensive and difficult to collect. Proxies inferred
from either scientific surveys or commercial catch data are often used to model
distributions, often with sparse data at limited spatial and temporal
resolution. \\

@@

We simulate 50 years of fishing and use the results from the fisheries catch to
draw inference on the underlying population structures. We compare this
inference to i) a simulated fixed-site sampling design commonly used for
fisheries monitoring purposes, and ii) the true underlying population
structures input to the simulation, to establish the potential and limitations
of fishery-dependent data - an inherently biased sampling method due to
fisher's targeting- to provide a robust picture of spatiotemporal
distributions. Finally, we simulate an area closure based on areas defined from
the known ("real-population") distribution, commercial catch data and survey
data at different temporal and spatial resolutions and assess their
effectiveness on reducing catches of a fish population. \\

@@

In order to understand \replaced[id=PD]{the consequences of using}{challenges
	that face} VMS-linked landings to draw inference on the underlying
population structure we develop a simulation model where population dynamics
are highly-resolved in space and time\replaced[id=PD]{. Being}{ and are}
known \added[id=PD]{directly} rather than inferred from sampling or commercial
catch\added[id=PD]{, we can use the population model to evaluate how inference from
	fisheries-dependent and fisheries independent sampling relates to the
	real population structure}. \replaced[id=PD]{In our model system
	p}{P}opulation movement is driven by random (diffusive) and directed
(advective) processes and we incorporate characterisation of a number of
different fishing \added[id=PD]{fleet dynamics} exploiting four fish populations
with different spatial and population demographics.\\

@@

\begin{enumerate}
	\item to understand how sampling-derived data reflects the underlying
		population structures. We compare at different spatial and
		temporal aggregations of data the real population to: 
		\begin{enumerate}
			\item the inferred population from a stratified fixed-site sampling survey design
				commonly used for fisheries monitoring purposes, otherwise know as a
				fisheries-independent survey,
			\item the inferred population from our fishery-dependent model which includes
				fishery-induced sampling dynamics.\\
		\end{enumerate}
	\item to understand the impact of data aggregation and source on
		spatial fisheries management measures we simulate a fishery closure to
		protect a species based on different spatial and temporal data
		aggregations:
		\begin{enumerate}
			\item as if the real spatial population structure were
				known,
			\item the fishery-independent inferred population
				structure
			\item the fishery-dependent inferred population
				structure
		\end{enumerate}
		We evaluate the theoretical "benefit" to the population of the
		closure(s), the effect on the other three populations and
		fishery catch.\todo{\added[id=JJ]{If the paper has two goals
				this should be clear from the start, but may be
				better over 2 MSs}\added[id=PD]{I would like to
				keep both parts, but have made clearer in how
				its set out. The closure scenarios form
				validation of the data aggregation, rather than
				effectiveness of the closures themselves - so
				its a continuation of the same question in my
				eyes}}\deleted[id=PD]{.  Further, we extend our
			analysis to a range of spatial and temporal scales to
			assess the impact of these processes on the success of
			the management measure.}\\
\end{enumerate}

@@

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics (fishing and natural mortality,
growth) operate on a daily time-step\replaced[id=PD]{, while p}{. P}opulation
movement occurs on a weekly time-step\replaced[id=PD]{. R}{, while r}ecruitment
\replaced[id=PD]{takes place}{occurs} periodcally each year for a set time
\replaced[id=PD]{duration}{period (e.g. 3 weeks)} specified for each
population\deleted[id=PD]{.}\added[id=PD]{, while the fishing module operates
	on a tow-by-tow basis (i.e. multiple events a day)}.  The simulation
framework is implemented in the statistical software package R
\citep{RCoreTeam2017} \replaced[id=PD]{and}{;} available as an R package from
the authors github site (\url{www.github.com/pdolder/MixFishSim}).\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Overview schematic  
\begin{figure}[!ht]
	\hspace{-4em}
\begin{tikzpicture}[%
    >=triangle 60,              % Nice arrows; your taste may be different
    start chain=going below,    % General flow is top-to-bottom
    node distance=8mm and 60mm, % Global setup of box spacing
    every join/.style={norm},   % Default linetype for connecting boxes
    scale=0.9,
    every node/.style={scale=0.8}]                 
    \label{fig:model}
    % ------------------------------------------------- 
% A few box styles 
% <on chain> *and* <on grid> reduce the need for manual relative
% positioning of nodes
\tikzset{
  base/.style={draw, on chain, on grid, align=center, minimum height=4ex},
  proc/.style={base, rectangle, text width=8em},
  test/.style={base, rectangle, text width=5em},
  term/.style={proc, rounded corners},
  % coord node style is used for placing corners of connecting lines
  coord/.style={coordinate, on chain, on grid, node distance=6mm and 25mm},
  % nmark node style is used for coordinate debugging marks
  nmark/.style={draw, cyan, circle, font={\sffamily\bfseries}},
  % -------------------------------------------------
  % Connector line styles for different parts of the diagram
  norm/.style={->, draw, lcnorm},
  free/.style={->, draw, lcfree},
  cong/.style={->, draw, lccong},
  it/.style={font={\small\itshape}}
}
% -------------------------------------------------
% Start by placing the nodes in the middle
\node [proc, densely dotted, it] (p0) {start};
% Use join to connect a node to the previous one 
\node [term, join, fill=lcnorm!25]      {Choose random location to fish};
\node [term, join, fill=lcfree!25] (p1){Sample catch};
\node [test, densely dotted] (t1) {if $(Rec)$};
\node [test, join, densely dotted] (t2) {if (Pop Movement)};
\node [test, join, densely dotted] (wk) {if (Pop Dynamics)};
\node [test, join, densely dotted] (t3) {if (Past Knowledge)};
\node [term, join, fill=lcnorm!25]  (p2) {Correlated Random Walk};
\node [term, join, fill=lcnorm!25]  (p3) {Choose new fishing location};
\node [proc, densely dotted, join, it] (p7) {end};

@@

% -------------------------------------------------
% A last flourish which breaks all the rules
\draw [->,purple, dotted, thick, shorten >=1mm]
  (p3.south) -- ++(5mm,-3mm)  -- ++(87mm,0mm) 
  |- node [black, near end, yshift=1.75em, it]
    {} ($(p1.north) + (0,0.3)$);
% -------------------------------------------------
\end{tikzpicture}
\caption{Overview Schematic of simulation model. The blue boxes indicate fleet
	dynamics processes, the green boxes population dynamics processes while
	the white boxes are the time steps at which processes occur; $t$ = tow,
	$tmax$ is the total number of tows; (Rec), (Pop Movement), (Pop
	Dynamics) logic gates for recruitment periods, population movement and
	population dynamics for each of the populations, (Past Knowledge) a
	switch whether to use a random (exploratory) or past knowledge
	(exploitation) fishing strategy.}
\label{fig:overview}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen as to discretise continuous population processes
	on a biologically relevant and computationally tractable timescale.}
\replaced[id=PD]{Under the population dynamics module}{Here,} population
biomass growth and depletion for pre-recruits and \deleted[id=PD]{fish}
recruited \added[id=PD]{fish} \deleted[id=PD]{to the fishery} are modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment. Biomass for each cell is incremented each day as
follows (the full parameter list is detailed in Table \ref{tab: pop_variable}):
\begin{equation*}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation*}
where $\rho$ is Brody's coefficient, shown to be approximately equal to
$e^{-K}$ when $K$ is the growth rate from a von Bertalanffy logistic growth
model \citep{Schnute1985}. $Wt_{R-1}$ is the weight of fish prior to
recruitment, while $Wt_{R}$ is the recruited weight. $\alpha_{d}$ represents
the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation*}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})}) *
B_{c,d}
\end{equation*}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. \todo{\added[id=CM]{[link $F$ to
	effort and catchability - as I think we have F as an emergent property
	of the fleets rather than something we solve for (I could be wrong
	though!) - catch for a vessel is a product of catchability and biomass,
	i.e. C = qB, but this catch is summed to solve for F. So its both
	really]}}\\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * B_{c,d})}{(\beta + B_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum, $B$ current SSB and
$\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations\replaced[id=JJ]{. An}{,
	with an} advection-diffusion process \deleted[id=JJ]{to}
control\added[id=JJ]{led} \deleted[id=JJ]{how the} population\deleted[id=JJ]{s}
move\replaced[id=JJ]{ment}{d} over time with a moving temperature covariate to
capture temporal dependencies\todo{\added[id=JJ]{What have a temperature
		covariate? Could just use time}\added[id=PD]{Was intended as
		some biological meaning - species thermal tolerances load onto
		the temperature effect - so could be different per species}}.
\deleted[id=JJ]{This was intended to balance realism in population movement,
	capturing the main directed and random processes, and practicality of
	modelling the population rather than individual fish.} \\

@@

The covariance structure affects the smoothness of the surfaces which the
process generates\replaced[id=PD]{;}{, and} we used the \textit{Matérn}
\deleted[id=PD]{family of} covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{as}{one where} the correlation strength weakens the further
the distance apart \deleted[id=PD]{(i.e. the correlation between $S(x)$ and
	$S(x')$ decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{The Matérn covariance structure models the spatial
	autocorrelation, where animal densities are observed to be more similar
	in nearby locations \citep{Tobler1970,
		F.Dormann2007}}\replaced[id=PD]{. It is}{The \textit{Matérn}
	correlation} is a two-parameter family where: 
\begin{center}
	$\rho(u) = \{2^{\kappa -
		1}\Gamma{\kappa}\}^{-1}(u/\phi)^{\kappa}K_{\kappa}(u/\phi)$
\end{center}
$K_{\kappa}(.)$ is a modified Bessel function of order $\kappa$, $\phi >
0$ is a scale parameter with the dimensions of distance, and $\kappa > 0$,
called the order, is a shape parameter which determines the smoothness of the
underlying process. \\

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, implementing different parameter settings to affect the
patchiness of the populations. Each population \replaced[id=PD]{was}{is}
initialised at a single location, and subsequently move\replaced[id=PD]{ed}{s}
according to a probabilistic distrib\added[id=PD]{u}tion based on habitat
suitability, temperature and distance from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$.\\

During specified weeks of the year, the habitat quality
\replaced[id=PD]{was}{is} modified for \added[id=PD]{user-defined} spawning
habitat\deleted[id=PD]{s}, \replaced[id=PD]{resulting in}{meaning} each
population \replaced[id=PD]{having}{has} a concentrated area where spawning takes
place and the population move\replaced[id=PD]{d}{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to
spawning.\todo{\added[id=JM]{What does it mean concisely? Areas are assigned?}
\added[id=PD]{Yes, the areas are pre-defined - I have amended to reflect and
	tried to clarify}} \\

The temperature field \replaced[id=PD]{was}{is} simulated to be on a gradient
from a South-Westerly to North-Easterly direction, with temperature in each
cell changing gradually on a week-by-week basis so that initially high
temperature areas cycle\added[id=PD]{d} to lower temperatures and low
temperature areas vice versa. Each population $p$ \replaced[id=PD]{wa}{i}s
assigned a thermal tolerance with mean, $\mu\added[id=PD]{_{p}}$ and variance,
$\sigma^2\added[id=PD]{_{p}}$ so that each cell and population temperature
suitability is defined that:
\begin{equation}
	Tol_{c, p\added[id=PD]{, wk}} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, \added[id=PD]{wk}}$ is the tolerance of population $p$ for
cell $c$ \added[id=PD]{in week $wk$}, $T_{c, \added[id=PD]{wk}}$ is the
temperature in the cell \added[id=PD]{given the week} and
$\mu\added[id=PD]{_{p}}$ and $\sigma^2\added[id=PD]{_{p}}$ the mean and
standard deviation of the population temperature tolerance. \\

\added[id=PD]{The final process resulted in independent populations structure
	and movement patterns, with population movement occuring on a weekly
	basis. This process approximated the demographic shifts in fish
	populations throughout a year with seasonal spawning patterns.}

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, \textit{Q}, and a population specific efficiency,
${Q_{p}}$.  Thus, the product of these parameters affect\replaced[id=PD]{s}{ed}
the overall catch rates for the fleet and the preferential targeting of one
population over another. This, in combination with the parameter choice for the
step-function \added[id=PD]{defined below} (as well as some randomness from the
exploratory fishing process) determine\replaced[id=PD]{d}{s} the preference of
fishing locations for the fleet.  All species prices
\replaced[id=PD]{were}{are} kept the same across fleets \replaced[id=PD]{and
	seasons}{, though can be made to vary seasonally}.

@@

\subsection{Population parametrisation}

We parametrised the simulation model for four populations with differing
habitat preference\replaced{,}{ and} temperature tolerances (Figures
\ref{S-fig:1}, \ref{S-fig:3}, \ref{S-fig:4}, \ref{S-fig:5}, \ref{S-fig:6},
\ref{S-fig:7}), population demographic and recruitment functions. In addition,
each of the populations has two defined spawning areas which result in the
populations moving towards these areas in given weeks (Figure \ref{S-fig:2})
and population-specific movement rates (Table \ref{tab:1}). The realised
movement of the populations for a number of weeks is shown in Figure
\ref{S-fig:9} while the realised daily fishing mortality are shown in Figure
\ref{S-fig:10}.

\subsection{Fleet parametrisation}
